On the characteristic and diameter of planar integral point sets \footnote{ The work of the first author was partially supported by the Theoretical Physics and Mathematics Advancement Foundation “BASIS”. }

### 2025-03-02 char-4k-bound/char-4k-bound.tex
[4k3] Unification: \overline{\mathfrak{M}_n}' -> \overline{\mathfrak{M}}'_n

\begin{lemma}
	\label{lem:no_distance_2}
	Every set $M \in \overline{\mathfrak{M}_n}'$ with an edge $|M_1M_2|=2$ has a cardinality $n=3$.
\end{lemma}

\begin{proof}
	Let $M \in \overline{\mathfrak{M}_n}'$ and $M_1, M_2 \in M$.
	Then all points in the set $M$ lie on $cr(M_1, M_2)$.
	Otherwise, according to Lemma~\ref{lem:pre_weeding_II}, $\operatorname{char}{M}={8k+7}$, $k\in \mathbb{N}_0$.

	Suppose, to the contrary, that $M={M_1, M_2, M_3, M_4} \in \overline{\mathfrak{M}_4}'$.
	Employ the Grid Theorem and set $M_1(-1;0), M_2(1;0), O(0;0)$.
	Then $M_3$ and $M_4$ lie on the perpendicular bisector to $M_1M_2$.
	Since the distance $|M_3M_4|$ is an integer, the area of triangle $M_1M_3M_4$ is a rational number.
	By Definition~\ref{def:char}, this means that $\operatorname{char}{M}={1}$.
	Then by the Grid Theorem we have $M_3(0;t/4)$. Without loss of generality, let us assume that $t\in\mathbb{N}$. Let $|M_1M_3|=s$. Applying the Pythagorean theorem to triangle $OM_1M_3$, we get
	$$
	1+\frac{t^2}{16}=s^2,
	$$
	or equivalently,
	\begin{equation}
	\label{eq: 42}
	16s^2-t^2=16.
	\end{equation}
	Now, we find the solutions to equation~\eqref{eq: 42} in positive integers.
	To do this, we introduce the substitution: $t=4k$, $k\in\mathbb{N}$.
	Then equation~\eqref{eq: 42} takes the form:
	\begin{equation}
	\label{eq: 43}
	s^2-k^2=1.
	\end{equation}
	Break down the left-hand side of equation~\eqref{eq: 43} into factors:
	\begin{equation}
	\label{eq: 44}
	(s-k)(s+k)=1.
	\end{equation}
	Since $s,k\in\mathbb{N}$, the number 1 can be represented as the product of two integers in two ways: $1\cdot1$ and $-1\cdot(-1)$. In the first case, $(s-k)=1$ and $(s+k)=1$, which gives $s=1,k=0$, which contradicts that $k\in\mathbb{N}$.
	In the second case, $(s-k)=-1$ and $(s+k)=-1$, which gives $s=-1$, $k=0$, which also contradicts that $s,k\in\mathbb{N}$.

@@

\begin{corollary}
	\label{cor:sqrt_5_3}
	In a triangle $ABC$ with sides $3 \leq a \leq b \leq c$, where ${A, B, C} \subset M \in \overline{\mathfrak{M}_4}'$, the smallest height is at least $(2a-1)^{1/2}$.
\end{corollary}

@@

\begin{lemma}
	\label{lem:intersection_enumeration}
	Let ${M_1, M_2, M_3, M_4} \subset M \in \overline{\mathfrak{M}_n}'$ (points $M_2$ and $M_3$ may coincide, while the others are distinct), where $n \ge 4$. Then $\# M \le |M_1 M_2| \cdot |M_3 M_4| - 2$.
\end{lemma}

@@

\begin{theorem}
	\label{thm:main_result}
	Let $M \in \overline{\mathfrak{M}}_n'$, i.e., $M$ is a set in semi-general position with characteristic different from $4k+3$, $k \in \mathbb{N}_0$.
	Then, for every integer $n \geq 3$, the inequality
	\begin{equation}
		\operatorname{diam} M \geq \left( \frac{25n}{36} \right)^{5/4}
	\end{equation}
	holds.
\end{theorem}

@@

	Let us consider $M \in \overline{\mathfrak{M}_n}'$, $n \geq 4$, $\operatorname{diam} M = p$.

>>>>

\begin{lemma}
	\label{lem:no_distance_2}
	Every set $M \in \overline{\mathfrak{M}}'_n$ with an edge $|M_1M_2|=2$ has the cardinality $n=3$.
\end{lemma}

\begin{proof}
	Let $M \in \overline{\mathfrak{M}}'_n$ and $M_1, M_2 \in M$.
	Then all points in the set $M$ lie on $cr(M_1, M_2)$.
	Otherwise, according to Lemma~\ref{lem:pre_weeding_II}, $\operatorname{char}{M}={8k+7}$, $k\in \mathbb{N}_0$.

	Suppose, to the contrary, that $M={M_1, M_2, M_3, M_4} \in \overline{\mathfrak{M}}'_4$.
	Employ the Grid Theorem and set $M_1(-1;0), M_2(1;0), O(0;0)$.
	Then $M_3$ and $M_4$ lie on the perpendicular bisector to $M_1M_2$.
	Since the distance $|M_3M_4|$ is an integer, the area of triangle $M_1M_3M_4$ is a rational number.
	By Definition~\ref{def:char}, this means that $\operatorname{char}{M}={1}$.
	Then by the Grid Theorem we have $M_3(0;t/4)$. Without loss of generality, let us assume that $t\in\mathbb{N}$. Let $|M_1M_3|=s$. Applying the Pythagorean theorem to triangle $OM_1M_3$, we get
	$$
	1+\frac{t^2}{16}=s^2,
	$$
	or equivalently,
	\begin{equation}
	\label{eq: 42}
	16s^2-t^2=16.
	\end{equation}
	Now, we find the solutions to equation~\eqref{eq: 42} in positive integers.
	To do this, we introduce the substitution: $t=4k$, $k\in\mathbb{N}$.
	Then equation~\eqref{eq: 42} takes the form:
	\begin{equation}
	\label{eq: 43}
	s^2-k^2=1.
	\end{equation}
	Break down the left-hand side of equation~\eqref{eq: 43} into factors:
	\begin{equation}
	\label{eq: 44}
	(s-k)(s+k)=1.
	\end{equation}
	Since $s,k\in\mathbb{N}$, the number 1 can be represented as the product of two integers in two ways: $1\cdot1$ and $-1\cdot(-1)$. In the first case, $(s-k)=1$ and $(s+k)=1$, which gives $s=1,k=0$, which contradicts that $k\in\mathbb{N}$.
	In the second case, $(s-k)=-1$ and $(s+k)=-1$, which gives $s=-1$, $k=0$, which also contradicts that $s,k\in\mathbb{N}$.

@@

\begin{corollary}
	\label{cor:sqrt_5_3}
	In a triangle $ABC$ with sides $3 \leq a \leq b \leq c$, where ${A, B, C} \subset M \in \overline{\mathfrak{M}}'_4$, the smallest height is at least $(2a-1)^{1/2}$.
\end{corollary}

@@

\begin{lemma}
	\label{lem:intersection_enumeration}
	Let ${M_1, M_2, M_3, M_4} \subset M \in \overline{\mathfrak{M}}'_n$ (points $M_2$ and $M_3$ may coincide, while the others are distinct), where $n \ge 4$. Then $\# M \le |M_1 M_2| \cdot |M_3 M_4| - 2$.
\end{lemma}

@@

\begin{theorem}
	\label{thm:main_result}
	Let $M \in \overline{\mathfrak{M}}'_n$, i.e., $M$ is a set in semi-general position with characteristic different from $4k+3$, $k \in \mathbb{N}_0$.
	Then, for every integer $n \geq 3$, the inequality
	\begin{equation}
		\operatorname{diam} M \geq \left( \frac{25n}{36} \right)^{5/4}
	\end{equation}
	holds.
\end{theorem}

@@

	Let us consider $M \in \overline{\mathfrak{M}}'_n$, $n \geq 4$, $\operatorname{diam} M = p$.
